---
title: "6306_Unit_4-3_Exercise"
author: "Karl J. Jurek"
date: "November 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ZillowR API

| 1. Reference Manual: ZillowR.pdf
|       https://cran.r-project.org/web/packages/ZillowR/ZillowR.pdf
| 2. Requirements:  
|       A. zws_id
|                   1) Register at: http://www.zillow.com/webservice/Registration.htm
|       B. zpid (depending on the request you make)
|                   EXAMPLE:  To get a zpid go to the Zillow website and select a property.  Then pull the zpid from the URL.
|                             https://www.zillow.com/homes/for_sale/Layton-UT/119633298_zpid/57030_rid/500000-_price/2070-_mp/globalrelevanceex_sort/41.123946,-111.89013,41.069096,-111.982398_rect/13_zm/0_mmm/
| 
| 3. Installation: 
|       A. install.packages("ZillowR", dependencies=TRUE)
|       B. library(ZillowR)
|       C. library(XML)

```{r installation, echo=TRUE}
#install.packages("ZillowR", dependencies=TRUE, repos="http://cran.us.r-project.org")
#library(ZillowR)
```

## EXAMPLE

| The following code was acquired from: https://stackoverflow.com/questions/38817153/zillow-api-with-r-xml-issue
| To make the code work there are a couple of things to note:
|       1. You will need to put in your own <b>zws_id</b>
|       2. If you change the address you will need to re-source the script code.

```{r script_window, echo=TRUE}
#This code will be placed in the Script window
#getValRange <- function(x, hilo) {
#  ifelse(hilo %in% unlist(dimnames(x)), x["text",hilo][[1]], NA)
#}

#out <- apply(zresults_list, MAR=2, function(property) {
#  zpid <- property$zpid
#  links <- unlist(property$links)
#  address <- unlist(property$address)
#  z <- property$zestimate
#  zestdf <- list(
#    amount=ifelse("text" %in% names(z$amount), z$amount$text, NA),
#    lastupdated=z$"last-updated",
#    valueChange=ifelse(length(z$valueChange)==0, NA, z$valueChange),
#    valueLow=getValRange(z$valuationRange, "low"),
#    valueHigh=getValRange(z$valuationRange, "high"),
#    percentile=z$percentile)  
#  list(id=zpid, links, address, zestdf)
#})

#zdata <- as.data.frame(do.call(rbind, lapply(out, unlist)), row.names=seq_len(length(out)))
#zdata
```

```{r console, echo=TRUE}
#Get search results
#zresults=GetSearchResults(address='411 E 6th St', citystatezip='10009', rentzestimate = FALSE, zws_id ="[YOUR_ZWS_ID]", url="http://www.zillow.com/webservice/GetSearchResults.htm")

#zresults_list<- xmlToList(zresults$response[["results"]])
#zdata <- as.data.frame(do.call(rbind, lapply(out, unlist)), row.names=seq_len(length(out)))
#write.csv(zdata, "NY_Zillow_Data.csv", na="NA")
```
