---
title: "KJurek_Livesession3assignment"
author: "Karl J. Jurek"
date: "November 24, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
GitHub Cloning (20 points): Using Git, clone the following GitHub repository to
your local machine: https://github.com/caesar0301/awesome-public-datasets. In
RMarkdown, please show the code (commented out, as it’s not R syntax) that you
used to create a new directory, navigate to the appropriate directory, and clone the
repository to it. One Git command per line, please.

|       <b>ANSWER:</b>
|
|               Starting from the C:/ directory
|               $ mkdir GIT_Repositories
|               $ cd git_repositories
|               $ mkdir 6306_Unit_3_HW
|               $ cd 6306_Unit_3_HW
|               $ git clone https://github.com/caesar0301/awesome-public-datasets
|
|
## Question 2: 
Data Summary (20 points): From this aforementioned cloned repo, please extract
titanic.csv.zip. To be clear, this does not have to be done in Git or command line.

|
## Question 2A: 
In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to
df. This dataset follows the passengers aboard the Titanic, including their fees
paid, rooms rented, and survivorship status.

|

```{r question_2A, echo=TRUE}
df=read.csv("C:\\GIT_Repositories\\6306_Unit_3_HW\\awesome-public-datasets\\Datasets\\titanic\\titanic.csv", header=TRUE, sep=",", stringsAsFactors=FALSE)
```

|
## Question 2B: 
Output the respective count of females and males aboard the Titanic. Plot the
frequency of females and males. Be sure to give an accurate title and label the axes.
```{r question_2B, echo=TRUE}
barplot(table(df$Sex), xlab="Passenger Gender", ylab="Count", names.arg=c("Females = 314", "Males = 577"), main="Gender Counts On The Titanic", col=3)
```

|
## Question 2C: 
Please use one apply function (to review: swirl() modules 11, 12) to output the
means of Age, Fare, and Survival. Make sure the output is a real number for
all three means.
```{r question_2C, echo=TRUE}
Age_Fare_Survived_mean=lapply(df[, c("Age", "Fare", "Survived")], mean, na.rm=TRUE)
Age_Fare_Survived_mean
```

## Question 3: 
Function Building (30 points): You research sleep and just got your first data set. Later,
you’ll have another dataset with the same column names, so you want to create a helper
function that you can analyze this dataset and the next. Load sleep_data_01.csv (found at
http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/index.html). Questions 3A through 3D
should be answered in function(x){}. 3E can be outside of the function.
```{r question_3, echo=TRUE}
sleep_data=(read.csv(url("http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv"), header=TRUE, sep=",", stringsAsFactors=FALSE))
```
|

```{r question_3_function, echo=TRUE}
ds_helper=function(type, dataset, col, rpt_obj, funk, df_flds, rnd)
{
  #################################################################################################
    #CREATED BY: Karl J. Jurek
    #CREATION DATE: 25 Nov 2018
    #LAST UPDATED: 25 Nov 2018
    #types are:
      #'data' - creates objects based on the function that is passed in
      #'report' - creates a simple report based on the previous created objects
    #dataset is the dataset that is to be used
    #col is the specific column to be accessed in the data set
    #funk is the function to be called
      #built in functions, such as 'median', can be passed individually or
      #if you want to pass in two or more functions to create a vector it can do that as well
        #EXAMPLE:  
          #rses_mean_and_sd=ds_helper(dataset=sleep_data, col="RSES", funk=c(mean,sd))
          #results in:  rses_mean_and_sd equal to [1] 18.114943  6.176522
    #report allows for three options using the 'funk' object
      #make - will create the report data field based on predetermined objects
        #fnk_report - the dataframe that will be passed back with the desired data
        #df_fields - the object's values desired in the report
      #rename - will rename the columns of the report based on the values in the 'df_fields' object 
        #rpt_obj - the report data frame that the names are going to be changes on
        #df_flds - the new column names      
      #round - will round the values in the report based on the value passed in the 'rnd' object
        #rpt_obj - the report data frame whose values are going to be rounded
        #rnd - the number of decimals to round the values in the report
  #################################################################################################
  
  if (type=="data")
  {
    #creates an object based on the length(size) of the values in the 'funk' variable
    result=c(length(funk))
    if (length(funk)<1)
    {
      #some error message - not added yet KJJ 11/25/2018
    }
    else
    {
      if (length(funk)==1)
      {
        #performs the function that is passed
        result=funk(dataset[, col], na.rm=TRUE)
      }
      else
      {
        #if the funk object has more than one function to be performed it will loop through each
        #function and place the results in one vector
        for (i in 1:length(funk))
        {
          result[i]=funk[[i]](dataset[, col], na.rm=TRUE)
        }
      }
      return(result)
    }
  }
  else if (type=="report")
  {
    #this will create a data frame based on the values of the objects passed
    if (funk=="make")
    {
      #creates a data frame based on the number of objects passed 
      fnk_report=data.frame(length(df_flds))
      
      #adds the values of the objects that were passed into the data frame
      for (i in 1:length(df_flds))
      {
        fnk_report[i]=data.frame(df_flds[i])
      }
      
      return(fnk_report)
      
    }
    else if (funk=="rename")
    {
      #renames the column names for the data frame that was passed in with the 
      #values passed into df_flds 
      names(rpt_obj)=df_flds
      
      return(rpt_obj)
    }
    else if (funk=="round")
    {
      
      #rounds any values of the passed data frame to the number of decimals indicated in 'rnd'
      rpt_obj=round(rpt_obj, digits=rnd)
      
      return(rpt_obj)
    }
  }
}

```

|
## Question 3A: 
Create objects for the median Age, the minimum and maximum Duration of sleep, and
the mean and standard deviation of the Rosenberg Self Esteem scale (RSES). You
may need to specify a few options like in Problem 2 and live session.

|
```{r question_3A, echo=TRUE}
age_median=ds_helper(type="data", dataset=sleep_data, col="Age", funk=median)
age_median

dur_min=ds_helper(type="data", dataset=sleep_data, col=15, funk=min)
dur_min

dur_max=ds_helper(type="data", dataset=sleep_data, col=15, funk=max)
dur_max

rses_mean=ds_helper(type="data", dataset=sleep_data, col="RSES", funk=mean)
rses_mean

rses_sd=ds_helper(type="data", dataset=sleep_data, col="RSES", funk=sd)
rses_sd

rses_mean_and_sd=ds_helper(type="data", dataset=sleep_data, col="RSES", funk=c(mean,sd))
rses_mean_and_sd
```

## Question 3B: 
Create a data.frame object called report: it should consist of the median age, the RSES
mean and standard deviation respectively divided by five (since there are five
questions and these scores are summed), and the range of Duration (the statistical
definition of range; it should be a single number.)

|
```{r question_3B, echo=TRUE}
report=ds_helper(type="report", funk="make", df_flds=c(age_median, rses_mean/5, rses_sd/5, dur_max-dur_min))
report
```

## Question 3C: 
Change the column names of this data.frame to MedianAge, SelfEsteem, SE_SD, and
DurationRange.

|
```{r question_3C, echo=TRUE}
new_names=c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange" )
report=ds_helper(type="report", funk="rename", rpt_obj=report, df_flds=new_names)
report
```

## Question 3D: 
Round the report to at most 2 digits: leave this as the closing line to the function.
|
```{r question_3D, echo=TRUE}
report_round1=ds_helper(type="report", funk="round", rpt_obj=report, rnd=1)
report_round1

report_round2=ds_helper(type="report", funk="round", rpt_obj=report, rnd=2)
report_round2
```

## Question 3E: 
Finally, run the function on your sleep data to show the output.

|       <b>ANSWER:</b>
|               See results from Questions 3A-3D.
|
## Question 4: 
FiveThirtyEight Data (30 points): Navigate on GitHub to
https://github.com/rudeboybert/fivethirtyeight and read README.md. It will include
everything you need.

## Question 4A:
Install the fivethirtyeight package.

```{r question_4A, echo=TRUE}
install.packages("fivethirtyeight", dependencies=TRUE, repos="http://cran.us.r-project.org")
library(fivethirtyeight)
data(package="fivethirtyeight")
```

|
## Question 4B:
In the listing of Data sets in package ‘fivethirtyeight,’ assign the 22nd data set to an
object ‘df538.’

```{r question_4B, echo=TRUE}
df538=college_recent_grads
head(df538)
```

|
## Question 4C:
Use a more detailed list of the data sets to write out the URL in a comment to the
related news story.

```{r question_4C, echo=TRUE}
install.packages('rvest', dependencies=TRUE, repos="http://cran.us.r-project.org")
library('rvest')
url='https://cran.r-project.org/web/packages/fivethirtyeight/vignettes/fivethirtyeight.html'
webpage=read_html(url)
#Select the destination using the 'Selector Gadget' tool
ref_html_nodes=html_nodes(webpage, ".uri")
article_links=html_text(ref_html_nodes)
print(article_links[22])
```

|
## Question 4D:
Using R command(s), give the dimensions and column names of this data frame.

```{r question_4D, echo=TRUE}
dim(df538)

#NOTE: The dimensions of the dataframe are also listed in the structure information
str(df538)
```

|
## Question 5:
Data Summary (30 points): Use your newly assigned data frame from question 4 for
this question.

|
## Question 5A1:
Write an R command that gives you the column names of the data frame.

```{r question_5A1, echo=TRUE}
#displays the column names in the 'df538' dataframe
colnames(df538)
```

|
## Question 5A2:
Right after that, write one that counts the number of columns but not rows. 
Hint: The number should match one of your numbers in Question 4d for dimensions.

```{r question_5A2, echo=TRUE}
#displays the count of the number of columns in the 'df538' dataframe
ncol(df538)
```

|
## Question 5B:
Generate a count of each unique major_category in the data frame. I recommend using
libraries to help. To be clear, this should look like a matrix or data frame containing
the major_category and the frequency it occurs in the dataset. Assign it to
major_count.

```{r question_5B, echo=TRUE}
#Gives then frequency of each value in the df538$major_category and places it in a dataframe called major_count
major_count=as.data.frame(table(df538$major_category))
#displays the dataframe 'major_count' data
major_count
```

|
## Question 5C:
To make things easier to read, put par(las=2) before your plot to make the text
perpendicular to the axis. Make a barplot of major_count. Make sure to label the title
with something informative (check the vignette if you need), label the x and y axis, and
make it any color other than grey. Assign the major_category labels to their respective
bar. Flip the barplot horizontally so that bars extend to the right, not upward. All of
these options can be done in a single pass of barplot(). Note: It’s okay if it’s wider than
the preview pane.

```{r question_5C, echo=TRUE}
#renaming the first column in the major_count object
names(major_count)=c("Majors", "Freq")
#ordering the Frequency from lowest to highest
new_order=major_count[order(major_count$Freq, decreasing=TRUE),]
#assigning the major names for the y-axis label of the bar plot
major_names=new_order$Majors
#setting the labels to horizontal
par(las=2)
#setting the margins
par(mar=c(5,14,2,1))
#assigning the title to be used in the barplot
title="Count of Sub-Categories of University Majors"
#assigning the x-axis label to be used in the barplot
x_label="Count"
#assigning the y-axis label if it was to be used in the barplot
y_label="Major Categories"
barplot(new_order$Freq, horiz=TRUE, names.arg=major_names, main=title, xlab=x_label, xlim=c(0,30), col=3)
```

|
## Question 5D:
Write the fivethirtyeight data to a csv file. Make sure that it does not have row labels.

```{r question_5D, echo=TRUE}
#Creates a .csv file with the information from the 'new_order' dataframe in the working directory
write.csv(new_order, file="Count_of_Major_Categories.csv", row.names=FALSE, na="NA")
```