---
title: "KJurek_Livesession5assignment"
author: "Karl J. Jurek"
date: "December 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Data Munging (30 points): 
Utilize yob2016.txt for this question. This file is a series of
popular children’s names born in the year 2016 in the United States. It consists of three
columns with a first name, a gender, and the amount of children given that name. However,
the data is raw and will need cleaning to make it tidy and usable.

### Question 1A: 


#### Question 1A-1:
First, import the .txt file into R so you can process it. Keep in mind this is not a CSV file. 
You might have to open the file to see what you’re dealing with.

```{r question_1A-1, echo=TRUE}
# Uploading the 'yob2016.txt' file from my GitHub directory into the 'df' data frame
df=read.delim("https://raw.githubusercontent.com/kjurekSMU/Homework/2018-11-26_HW_Start/6306_HW/6306_Unit_5_HW/yob2016.txt", header=FALSE, sep=";", stringsAsFactors=FALSE)
```

#### Question 1A-2:     
Assign the resulting data frame to an object, df, that consists of three columns with human readable column names for each.

```{r question_1A-2, echo=TRUE}
# Assigning useful names to the columns in the 'df' data frame
names(df)=c("Name", "Gender", "Count")
```

### Question 1B:
Display the summary and structure of df
```{r question_1B, echo=TRUE}
# Viewing the summary of the data of the 'df' dataframe
summary(df)
# Validating the structure of the 'df' dataframe
str(df)
```

### Question 1C:
Your client tells you that there is a problem with the raw file. One name was entered twice and misspelled. 
The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally 
put three y’s at the end of the name. Write an R command to figure out which name it is and display it.
```{r question_1C, echo=TRUE}
# Looking for any name with three (3) y's [e.g., 'yyy'] in it 
df[grepl("yyy", df$Name),]
```


### Question 1D:
Upon finding the misspelled name, please remove this particular observation, as the client says it’s redundant. 
Save the remaining dataset as an object: y2016
```{r question_1D, echo=TRUE}
# Removing offending row, 212, and placing all other data in the 'y2016' data frame
y2016=df[-212,]
# Validating the structure of the new data frame 'y2016'
str(y2016)
```

## Question 2: Data Merging (30 points): 
Utilize yob2015.txt for this question. This file is similar to yob2016, but contains names, gender, 
and total children given that name for the year 2015.

### Question 2A: 
Like 1A, please import the .txt file into R. Look at the file before you do. 
You might have to change some options to import it properly. 
Again, please give the dataframe human-readable column names. Assign the dataframe to y2015.

```{r question_2A, echo=TRUE}
# Uploading the 'yob2016.txt' file from my GitHub directory into the 'df' data frame
y2015=read.delim("https://raw.githubusercontent.com/kjurekSMU/Homework/2018-11-26_HW_Start/6306_HW/6306_Unit_5_HW/yob2015.txt", header=FALSE, sep=",", stringsAsFactors=FALSE)
# Assigning useful names to the columns in the 'df' data frame
names(y2015)=c("Name", "Gender", "Count")
# Validating the structure of the new data frame 'y2015'
str(y2015)
```

### Question 2B: 

#### Question 2B-1:
Display the last ten rows in the dataframe. 

```{r question_2B-1, echo=TRUE}
tail(y2015, 10)
```

#### Question 2B-2:
Describe something you find interesting about these 10 rows.

|       <b> ANSWER: </b>
|               The interesting thing about the data is that the smallest number of children with a particular name stops at a count of five (5)

### Question 2C: 
Merge y2016 and y2015 by your Name column; assign it to <b>final</b>. 
The client only cares about names that have data for both 2016 and 2015; 
there should be no NA values in either of your amount of children rows after merging.
```{r question_2C, echo=TRUE}
# Merging the 'y2016' and 'y2015' data frames on both name and gender
final=merge(y2016, y2015, by.x=c("Name", "Gender"), by.y=c("Name", "Gender"))
```

## Question 3: Data Summary (30 points): 
Utilize your data frame object <b>final</b> for this part.

### Question 3A: 

#### Question 3A-1:
Create a new column called “Total” in final that adds the amount of children in 2015 and 2016 together. 
```{r question_3A-1, echo=TRUE}
# Adding the 'y2016' count and 'y2015' count columns together in the 'Total' column
final$Total=final$Count.x + final$Count.y
# Assigning useful names to the columns in the 'final' data frame
names(final)[c(3,4)]=c("Count_2016", "Count_2015")
# Validating the structure of the new data frame 'final'
str(final)
```

#### Question 3A-2:
In those two years combined, how many people were given popular names?
```{r question_3A-2, echo=TRUE}
# Adding all of column 'Total' together
sum(final$Total)
```

### Question 3B: 

#### Question 3B-1:
Sort the data by Total. 
```{r question_3B-1, echo=TRUE}
# Sorting the data using the 'Total' column in descending order
final=final[order(-final$Total),]
```

#### Question 3B-2:
What are the top 10 most popular names?
```{r question_3B-2, echo=TRUE}
# Showing the top 10 most popular name
head(final, 10)
```

### Question 3C: 
The client is expecting a girl! 

#### Question 3C-1:
Omit boys
```{r question_3C-1, echo=TRUE}
# Sorting the data using the 'Total' column in descending order
girls=final[grepl("F", final$Gender),]
```

### Question 3C-2:
Give the top 10 most popular girl’s
```{r question_3C-2, echo=TRUE}
# Showing the top 10 most popular girl names
head(girls, 10)
```

### Question 3D:
Write these top 10 girl names and their Totals to a CSV file. Leave out the other columns entirely.
```{r question_3D, echo=TRUE}
# Writing only the 'Name' and 'Total' columns to a csv file
write.csv(girls[1:10, c(1,5)], file="Top_10_Girl_Names.csv", row.names=FALSE)
```

## Question 4: Upload to GitHub (10 points):
Push at minimum your RMarkdown for this homework assignment and a Codebook to one of your GitHub 
repositories (you might place this in a Homework repo like last week). The Codebook should contain
a short definition of each object you create, and if creating multiple files, which file it is
contained in. You are welcome and encouraged to add other files—just make sure you have a description 
and directions that are helpful for the grader.

|       <b> ANSWER: </b>
|               See GitHub directory:  https://github.com/kjurekSMU/Homework/tree/2018-11-26_HW_Start/6306_HW/6306_Unit_5_HW
|
|
|
|
|
|
|
|
| #######END OF DOCUMENT#######